---
title: "Automated Machine Learning with H2O (I)"
author: "Ana Jade"
---

Employee attrition is defined as employee quitting. The goal of this data visualization is to try and see if you can predict if an employee is about to quit.

# Load libraries
```{r}
library(tidyverse)
library(readxl)
library(skimr)
library(GGally)

```

# Load data
```{r}
employee_attrition_tbl <- read_csv("data/04_data/datasets-1067-1925-WA_Fn-UseC_-HR-Employee-Attrition.csv")
employee_attrition_tbl
```
Now get the definitions:
```{r}
path_data_definitions <- "data/04_data/data_definitions.xlsx"
definitions_raw_tbl   <- read_excel(path_data_definitions, sheet = 1, col_names = FALSE)
```

# View link between different attributes

First, create a function to plot the ggpairs.
```{r}
plot_ggpairs <- function(data, color = NULL, density_alpha = 0.5) {
  
  color_expr <- enquo(color)
  
  if (rlang::quo_is_null(color_expr)) {
    
    g <- data %>%
      ggpairs(lower = "blank") 
    
  } else {
    
    color_name <- quo_name(color_expr)
    
    g <- data %>%
      ggpairs(mapping = aes_string(color = color_name), 
              lower = "blank", legend = 1,
              diag = list(continuous = wrap("densityDiag", 
                                            alpha = density_alpha))) +
      theme(legend.position = "bottom")
  }
  
  return(g)
  
}
```
## Compensation features
The following figure presents the interaction between compensation features and attrition.
```{r, fig.width=10, fig.height=7}
employee_attrition_tbl %>%
  select(Attrition, contains("income"), contains("rate"), contains("salary"), contains("stock")) %>%
  plot_ggpairs(Attrition)
```
**1 - What can you deduce about the interaction between Monthly Income and Attrition?**
c. Those that are leaving have a lower Monthly Income 

**2 - What can you deduce about the interaction between Percent Salary Hike and Attrition?**
d. It's difficult to deduce anything based on the visualization

**3 - What can you deduce about the interaction between Stock Option Level and Attrition?**
c. It's difficult to deduce anything based on the visualization

## Survey Results
The following figure presents the interaction between survey results and attrition.
```{r, fig.width=10, fig.height=7}
employee_attrition_tbl %>%
  select(Attrition, contains("satisfaction"), contains("life")) %>%
  plot_ggpairs(Attrition)
```

**4 - What can you deduce about the interaction between Environment Satisfaction and Attrition?**
a. A higher proportion of those leaving have a low environment satisfaction level

**5 - What can you deduce about the interaction between Work Life Balance and Attrition**
b. Those that are staying have a higher density of 2's and 3's

## Performance data
The following figure presents the interaction between performance data and attrition.
```{r, fig.width=10, fig.height=7}
employee_attrition_tbl %>%
  select(Attrition, contains("performance"), contains("involvement")) %>%
  plot_ggpairs(Attrition)
```

**6 - What Can you deduce about the interaction between Job Involvement and Attrition?**
a. Those that are leaving have a lower density of 3's and 4's

## Work-Life Features
The following figure presents the interaction between work-life features and attrition.
```{r, fig.width=10, fig.height=7}
employee_attrition_tbl %>%
  select(Attrition, contains("overtime"), contains("travel")) %>%
  plot_ggpairs(Attrition)
```

**7 - What can you deduce about the interaction between Over Time and Attrition?**
b. The proportion of those staying that are working Over Time are high compared to those that are not staying

## Training and Education
The following figure presents the interaction between training and education,  and attrition.
```{r, fig.width=10, fig.height=7}
employee_attrition_tbl %>%
  select(Attrition, contains("training"), contains("education")) %>%
  plot_ggpairs(Attrition)
```

**8 - What can you deduce about the interaction between Training Times Last Year and Attrition?**
b. People that leave tend to have less annual trainings

## Time-based features
The following figure presents the interaction between time-based features  and attrition.
```{r, fig.width=10, fig.height=7}
employee_attrition_tbl %>%
  select(Attrition, contains("years")) %>%
  plot_ggpairs(Attrition)
```

**9 - What can you deduce about the interaction between Years At Company and Attrition?**
b. People that leave tend to have less working years at the company

**10 - What can you deduce about the interaction between Years Since Last Promotion and Attrition?**
b. Those that are leaving have fewer years since last promotion than those that are staying










